---
title: "limpa analysis on mixed-species data"
format: html
---

# Summary

Differential expression (DE) analysis on the 3-organism mixture data from LFQbench (https://ftp.pride.ebi.ac.uk/pride/data/archive/2016/09/PXD002952/). 
We will look at the HYE110 dataset (TripleTOF 6600; 64-variable-window acquisition) processed by DIA-NN under the library-based mode, published in the DIA-NN 2020 paper. 
DIA-NN processed data were downloaded from https://osf.io/6g3ux/files/osfstorage. 
This file only includes a simple run through of the limpa pipeline, with minimal data exploration and quality assessment (this is a good quality dataset). 



# (Install and) Load packages

```{r}
#| eval: false

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("limpa")
```


```{r}
library(limpa)
```


# Data

```{r}
load("data/250510-mixedSpecies_cleaned.rda")
```



# limpa workflow

## Fit DPC

```{r}
dpcfit <- dpc(dat)
plotDPC(dpcfit)
dpcfit$dpc
```

## Protein quantification by DPC-Quant

```{r}
pt <- dpcQuant(dat, dpc = dpcfit)
dim(pt)
```


## DE analysis

```{r}
design <- model.matrix(~ 0 + Group, data = dat$samples)
contr <- makeContrasts(BvsA = GroupB - GroupA, levels = design)

fit <- dpcDE(pt, design)
fit <- contrasts.fit(fit, contrasts = contr)
fit <- eBayes(fit, robust = TRUE)

summary(decideTests(fit))
```


Mean-difference plot

```{r}
species.col <- c("blue", "red", "black")

plotMD(fit, coef = "BvsA", 
       status = as.factor(pt$genes$Species), 
       hl.col = species.col, 
       hl.cex = 0.5, 
       main = "Mean-difference plot (B vs A)")
abline(h = log2(3/30), col = "blue")
abline(h = 0, col = "black")
abline(h = log2(30/3), col = "red")
```



















# Session informaiton

```{r}
sessionInfo()
```



